---
title: "FinalProject"
author: "Madeline Egbert"
date: "4/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(lubridate)
```


```{r, error = FALSE, message = FALSE, warning = FALSE}
deathsByAge <- readr::read_csv('deaths_by_age_group.csv')

casesByAge <- readr::read_csv('cases_by_age_group.csv')

covid19Data <- readr::read_csv('covid19_vaccinations_in_the_united_states.csv')

casesdeathsByState <- readr::read_csv('united_states_covid19_cases_and_deaths_by_state.csv')
```


```{r}
deathsByAge
casesByAge
covid19Data
```

```{r, echo = FALSE}
covid19Data <- covid19Data %>%
  rename(
    State = `State/Territory/Federal Entity`,
    total.Doses = `Total Doses Delivered`,
    doses.Per.100K = `Doses Delivered per 100K`,
    doses.Per.100K.18plus = `18+ Doses Delivered per 100K`,
    total.Doses.Administered = `Total Doses Administered by State where Administered`,
    doses.Administered.Per.100k = `Doses Administered per 100k by State where Administered`,
    doses.Administered.18plus = `18+ Doses Administered by State where Administered`,
    doses.Administered.18plus.Per.100k = `18+ Doses Administered per 100K by State where Administered`,
    people.at.Least.One.Dose = `People with at least One Dose by State of Residence`,
    percent.at.Least.One.Dose = `Percent of Total Pop with at least One Dose by State of Residence`,
    people.at.Least.One.Dose.18plus = `People 18+ with at least One Dose by State of Residence`,
    percent.at.Least.One.Dose.18plus = `Percent of 18+ Pop with at least One Dose by State of Residence`,
    people.Fully.Vaccinated = `People Fully Vaccinated by State of Residence`,
    percent.Fully.Vaccinated = `Percent of Total Pop Fully Vaccinated by State of Residence`,
    people.Fully.Vaccinated.18plus = `People 18+ Fully Vaccinated by State of Residence`,
    percent.Fully.Vaccinated.18plus = `Percent of 18+ Pop Fully Vaccinated by State of Residence`,
    doses.Delivered.Pfizer = `Total Number of Pfizer doses delivered`,
    doses.Delivered.Moderna = `Total Number of Moderna doses delivered`,
    doses.Delivered.Janssen = `Total Number of Janssen doses delivered`,
    doses.Delivered.Unknown = `Total Number of doses from unknown manufacturer delivered`,
    doses.Administered.Janssen = `Total Number of Janssen doses administered`,
    doses.Administered.Moderna = `Total Number of Moderna doses administered`,
    doses.Administered.Pfizer = `Total Number of Pfizer doses adminstered`,
    doses.Administered.Unknown = `Total Number of doses from unknown manufacturer administered`,
    people.Fully.Vaccinated.Moderna = `People Fully Vaccinated Moderna Resident`,
    people.Fully.Vaccinated.Pfizer = `People Fully Vaccinated Pfizer Resident`,
    people.Fully.Vaccinated.Janssen = `People Fully Vaccinated Janssen Resident`,
    people.Fully.Vaccinated.Unknown = `People Fully Vaccinated Unknown 2-dose manufacturer Resident`,
    people.Fully.Vaccinated.Moderna.18plus = `People 18+ Fully Vaccinated Moderna Resident`,
    people.Fully.Vaccinated.Pfizer.18plus = `People 18+ Fully Vaccinated Pfizer Resident`,
    people.Fully.Vaccinated.Janssen.18plus = `People 18+ Fully Vaccinated Janssen Resident`,
    people.Fully.Vaccinated.Unknown.18plus = `People 18+ Fully Vaccinated Unknown 2-dose manufacturer Resident`,
    people.Two.Doses = `People with 2 Doses by State of Residence`,
    percent.1plus.Doses = `Percent of Total Pop with 1+ Doses by State of Residence`,
    Oneplus.Doses.18plus = `People 18+ with 1+ Doses by State of Residence`,
    percent.1plus.Doses.18plus = `Percent of 18+ Pop with 1+ Doses by State of Residence`,
    percent.2Doses = `Percent of Total Pop with 2 Doses by State of Residence`,
    people.18plus.2Doses = `People 18+ with 2 Doses by State of Residence`,
    percent.18plus.2Doses = `Percent of 18+ Pop with 2 Doses by State of Residence`,
    people.1plus.Doses = `People with 1+ Doses by State of Residence`,
    people.1plus.Doses.65plus = `People 65+ with at least One Dose by State of Residence`,
    percent.1plus.Doses.65plus = `Percent of 65+ Pop with at least One Dose by State of Residence`,
    people.Fully.Vaccinated.65plus = `People 65+ Fully Vaccinated by State of Residence`,
    percent.Fully.Vaccinated.65plus = `Percent of 65+ Pop Fully Vaccinated by State of Residence`,
    people.Fully.Vaccinated.Moderna.65plus = `People 65+ Fully Vaccinated_Moderna_Resident`,
    people.Fully.Vaccinated.Pfizer.65plus = `People 65+ Fully Vaccinated_Pfizer_Resident`,
    people.Fully.Vaccinated.Janssen.65plus = `People 65+ Fully Vaccinated_Janssen_Resident`,
    people.Fully.Vaccinated.Unknown.65plus = `People 65+ Fully Vaccinated_Unknown 2-dose Manuf_Resident`,
    doses.Administered.65plus = `65+ Doses Administered by State where Administered`,
    doses.Administered.Per.100K.65plus = `Doses Administered per 100k of 65+ pop by State where Administered`,
    doses.Delivered.Per.100K.65plus = `Doses Delivered per 100k of 65+ pop`
  )
```

```{r, error=FALSE, message=FALSE, warning=FALSE, echo = FALSE}
covid19Data$doses.Per.100K <- as.numeric(covid19Data$doses.Per.100K)
covid19Data$doses.Per.100K.18plus <- as.numeric(covid19Data$doses.Per.100K.18plus)
covid19Data$doses.Administered.Per.100k <- as.numeric(covid19Data$doses.Administered.Per.100k)
covid19Data$doses.Administered.18plus.Per.100k <- as.numeric(covid19Data$doses.Administered.18plus.Per.100k)
covid19Data$percent.at.Least.One.Dose <- as.numeric(covid19Data$percent.at.Least.One.Dose)
covid19Data$percent.at.Least.One.Dose.18plus <- as.numeric(covid19Data$percent.at.Least.One.Dose.18plus)
covid19Data$percent.Fully.Vaccinated <- as.numeric(covid19Data$percent.Fully.Vaccinated)
covid19Data$percent.Fully.Vaccinated.18plus <- as.numeric(covid19Data$percent.Fully.Vaccinated.18plus)
covid19Data$percent.1plus.Doses <- as.numeric(covid19Data$percent.1plus.Doses)
covid19Data$percent.1plus.Doses.18plus <- as.numeric(covid19Data$percent.1plus.Doses.18plus)
```

```{r}
str(covid19Data)
```

#What states had the highest increase in cases over the span of a week?


```{r}
mostcases7days <- casesdeathsByState %>%
   rename(State = `State/Territory`, ) %>%
  filter(State != 'United States of America', State != 'New York City') %>%
  select(State, `Cases in Last 7 Days`) %>%
  top_n(10, `Cases in Last 7 Days`)

str(mostcases7days)
```
```{r}
casesweek <- mostcases7days %>%
  group_by(State) %>%
  summarize(`Cases in Last 7 Days` = max(`Cases in Last 7 Days`)) %>%
  arrange(desc(`Cases in Last 7 Days`)) %>%
  head(n=5)

head(casesweek)
```

#explanation: The top 5 states with the most cases in the span of a week are as follows: Michigan, Florida, Pennsylvania, New Jersey and New York. Michigan has 10,164 more cases in a week than Florida which is decently large spike in Michigan. 

```{r}
topcases7days <- ggplot(mostcases7days, aes(x=State, y=`Cases in Last 7 Days`)) + geom_point() + ggtitle('Top States with Most Cases in Last 7 Days') + xlab('State') + ylab('Cases in Last 7 Days')

topcases7days + geom_abline(intercept = 20000, slope = 100, color = "red")
```

```{r}
CasesCount <- ggplot(mostcases7days, aes(x=State, y= `Cases in Last 7 Days`)) + geom_point() + ggtitle('Top States with Most Cases in Last 7 Days') + xlab('State') + ylab('Cases in Last 7 Days')

CasesCount + geom_abline(intercept = 100, slope = 50, color = "Purple")

ggplot(mostcases7days, aes(x= State, weight = `Cases in Last 7 Days`)) + geom_bar()
```




```{r}
ggplot(mostcases7days, aes(x=State, weight = `Cases in Last 7 Days`)) + geom_bar(fill = '#D23369') + scale_x_discrete(limit = c('Michigan', 'Florida', 'Pennsylvania', 'New Jersey', 'New York')) + ggtitle('Top 5 States with Most Cases in a Week') + xlab('State') + ylab('Cases in a Week')
```

#explanation: according to the results of the above graph,the top 5 states with the most cases recorded within the span of a week include: Michigan, Florida, Pennsylvania, New Jersey, and New York. I am suprised to not see California in the top 5 because of how large the state is and how populated it is as well. 



```{r}
ggplot(mostcases7days, aes(x = State, weight = `Cases in Last 7 Days`)) + geom_bar(fill = '#D23369') + scale_x_discrete(limit = c('Michigan', 'Florida', 'Pennsylvania', 'New Jersey', 'New York', 'Illinois', 'Texas', 'California', 'Ohio', 'Massachusetts')) + ggtitle('Top 10 States with the Most Cases in a Week') + xlab('State') + ylab('Cases in a Week') + coord_flip()
```

#explanation: according to the results of the above graph, we can see that Michigan takes the lead by a fair amount for having the highest count of cases in the last 7 days. Florida follows close behind with between 30,000-40,000 cases in the last 7 days. I was shocked to see New York and California farther away from the most amount of cases in the last week because they are such populated states that have had a really hard time controlling Covid within their state.

