---
title: "ds202_finalproject"
author: "Emily Pollock"
date: "4/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(lubridate)
library(tidyverse)
library(dplyr)
```

```{r, error=FALSE, message=FALSE, warning=FALSE}
deathsByAge <- readr::read_csv('deaths_by_age_group.csv')

casesByAge <- readr::read_csv('cases_by_age_group.csv')

covid19Data <- readr::read_csv('covid19_vaccinations_in_the_united_states.csv')

casesdeathsstates <- readr::read_csv('united_states_covid19_cases_and_deaths_by_state.csv')

deathsByAge
casesByAge
covid19Data
casesdeathsstates
```

```{r, echo = FALSE}
covid19Data <- covid19Data %>%
  rename(
    State = `State/Territory/Federal Entity`,
    total.Doses = `Total Doses Delivered`,
    doses.Per.100K = `Doses Delivered per 100K`,
    doses.Per.100K.18plus = `18+ Doses Delivered per 100K`,
    total.Doses.Administered = `Total Doses Administered by State where Administered`,
    doses.Administered.Per.100k = `Doses Administered per 100k by State where Administered`,
    doses.Administered.18plus = `18+ Doses Administered by State where Administered`,
    doses.Administered.18plus.Per.100k = `18+ Doses Administered per 100K by State where Administered`,
    people.at.Least.One.Dose = `People with at least One Dose by State of Residence`,
    percent.at.Least.One.Dose = `Percent of Total Pop with at least One Dose by State of Residence`,
    people.at.Least.One.Dose.18plus = `People 18+ with at least One Dose by State of Residence`,
    percent.at.Least.One.Dose.18plus = `Percent of 18+ Pop with at least One Dose by State of Residence`,
    people.Fully.Vaccinated = `People Fully Vaccinated by State of Residence`,
    percent.Fully.Vaccinated = `Percent of Total Pop Fully Vaccinated by State of Residence`,
    people.Fully.Vaccinated.18plus = `People 18+ Fully Vaccinated by State of Residence`,
    percent.Fully.Vaccinated.18plus = `Percent of 18+ Pop Fully Vaccinated by State of Residence`,
    doses.Delivered.Pfizer = `Total Number of Pfizer doses delivered`,
    doses.Delivered.Moderna = `Total Number of Moderna doses delivered`,
    doses.Delivered.Janssen = `Total Number of Janssen doses delivered`,
    doses.Delivered.Unknown = `Total Number of doses from unknown manufacturer delivered`,
    doses.Administered.Janssen = `Total Number of Janssen doses administered`,
    doses.Administered.Moderna = `Total Number of Moderna doses administered`,
    doses.Administered.Pfizer = `Total Number of Pfizer doses adminstered`,
    doses.Administered.Unknown = `Total Number of doses from unknown manufacturer administered`,
    people.Fully.Vaccinated.Moderna = `People Fully Vaccinated Moderna Resident`,
    people.Fully.Vaccinated.Pfizer = `People Fully Vaccinated Pfizer Resident`,
    people.Fully.Vaccinated.Janssen = `People Fully Vaccinated Janssen Resident`,
    people.Fully.Vaccinated.Unknown = `People Fully Vaccinated Unknown 2-dose manufacturer Resident`,
    people.Fully.Vaccinated.Moderna.18plus = `People 18+ Fully Vaccinated Moderna Resident`,
    people.Fully.Vaccinated.Pfizer.18plus = `People 18+ Fully Vaccinated Pfizer Resident`,
    people.Fully.Vaccinated.Janssen.18plus = `People 18+ Fully Vaccinated Janssen Resident`,
    people.Fully.Vaccinated.Unknown.18plus = `People 18+ Fully Vaccinated Unknown 2-dose manufacturer Resident`,
    people.Two.Doses = `People with 2 Doses by State of Residence`,
    percent.1plus.Doses = `Percent of Total Pop with 1+ Doses by State of Residence`,
    Oneplus.Doses.18plus = `People 18+ with 1+ Doses by State of Residence`,
    percent.1plus.Doses.18plus = `Percent of 18+ Pop with 1+ Doses by State of Residence`,
    percent.2Doses = `Percent of Total Pop with 2 Doses by State of Residence`,
    people.18plus.2Doses = `People 18+ with 2 Doses by State of Residence`,
    percent.18plus.2Doses = `Percent of 18+ Pop with 2 Doses by State of Residence`,
    people.1plus.Doses = `People with 1+ Doses by State of Residence`,
    people.1plus.Doses.65plus = `People 65+ with at least One Dose by State of Residence`,
    percent.1plus.Doses.65plus = `Percent of 65+ Pop with at least One Dose by State of Residence`,
    people.Fully.Vaccinated.65plus = `People 65+ Fully Vaccinated by State of Residence`,
    percent.Fully.Vaccinated.65plus = `Percent of 65+ Pop Fully Vaccinated by State of Residence`,
    people.Fully.Vaccinated.Moderna.65plus = `People 65+ Fully Vaccinated_Moderna_Resident`,
    people.Fully.Vaccinated.Pfizer.65plus = `People 65+ Fully Vaccinated_Pfizer_Resident`,
    people.Fully.Vaccinated.Janssen.65plus = `People 65+ Fully Vaccinated_Janssen_Resident`,
    people.Fully.Vaccinated.Unknown.65plus = `People 65+ Fully Vaccinated_Unknown 2-dose Manuf_Resident`,
    doses.Administered.65plus = `65+ Doses Administered by State where Administered`,
    doses.Administered.Per.100K.65plus = `Doses Administered per 100k of 65+ pop by State where Administered`,
    doses.Delivered.Per.100K.65plus = `Doses Delivered per 100k of 65+ pop`)
```

```{r, error=FALSE, message=FALSE, warning=FALSE, echo = FALSE}
covid19Data$doses.Per.100K <- as.numeric(covid19Data$doses.Per.100K)
covid19Data$doses.Per.100K.18plus <- as.numeric(covid19Data$doses.Per.100K.18plus)
covid19Data$doses.Administered.Per.100k <- as.numeric(covid19Data$doses.Administered.Per.100k)
covid19Data$doses.Administered.18plus.Per.100k <- as.numeric(covid19Data$doses.Administered.18plus.Per.100k)
covid19Data$percent.at.Least.One.Dose <- as.numeric(covid19Data$percent.at.Least.One.Dose)
covid19Data$percent.at.Least.One.Dose.18plus <- as.numeric(covid19Data$percent.at.Least.One.Dose.18plus)
covid19Data$percent.Fully.Vaccinated <- as.numeric(covid19Data$percent.Fully.Vaccinated)
covid19Data$percent.Fully.Vaccinated.18plus <- as.numeric(covid19Data$percent.Fully.Vaccinated.18plus)
covid19Data$percent.1plus.Doses <- as.numeric(covid19Data$percent.1plus.Doses)
covid19Data$percent.1plus.Doses.18plus <- as.numeric(covid19Data$percent.1plus.Doses.18plus)
```


#Which states have the highest amount of cases (pick top 5)? Do these states also administer the most vaccinations? (use cases and vaccines data) Which vaccine is the most popular for fully vaccinated people in these states?


```{r}
casesdeathsstates1 <- casesdeathsstates %>%
  rename(State = `State/Territory`, ) %>%
  select(State, `Cases in Last 7 Days`) %>%
  top_n(6, `Cases in Last 7 Days`)
```

```{r}
casesdeathsstates2 <- casesdeathsstates %>%
  rename(State = `State/Territory`, ) %>%
  select(State, `Total Cases`) %>%
  top_n(6, `Total Cases`)
```


```{r}
covid19Data1 <- covid19Data %>%
  select(State, total.Doses.Administered) %>%
  top_n(5, total.Doses.Administered)
```

```{r}
casesdeathsstates1
casesdeathsstates2
covid19Data1
```

#It makes sense that these five states have the highest total of administered doses because they make up the top five most populated states. Out of these states, Florida, New York and Pennsylvania also have some of the highest numbers for COVID-19 cases over the course of seven days. All of these five states have over 25,000 cases. And, when looking at total COVID-19 cases for these states, California, Florida and Texas have some of the highest totals, with m.

```{r}
covid19Data2 <- covid19Data %>%
  filter(State != 'American Samoa', State != 'Bureau of Prisons', State != 'District of Columbia', State != 'Dept of Defense', State != 'Federated States of Micronesia', State != 'Guam', State != 'Indian Health Svc', State != 'Marshall Islands', State != 'Northern Mariana Islands', State != 'Puerto Rico', State != 'Republic of Palau', State != 'Veterans Health', State != 'Virgin Islands')
covid19Data2
```

```{r}
casesdeathsstates3 <- casesdeathsstates %>%
  rename(State = `State/Territory`, ) %>%
  filter(State != 'American Samoa', State != 'District of Columbia', State != 'Federated States of Micronesia', State != 'Guam', State != 'Northern Mariana Islands', State != 'New York City', State != 'Puerto Rico', State != 'Palau', State != 'Republic of Marshall Islands', State != 'Virgin Islands', State != 'United States of America')
casesdeathsstates3
```

```{r}
dosesAdministered <- covid19Data2 %>%
  select(State, doses.Administered.Moderna, doses.Administered.Janssen, doses.Administered.Pfizer,
  doses.Administered.Unknown) %>%
  rename(
    moderna = doses.Administered.Moderna,
    pfizer = doses.Administered.Pfizer,
    janssen = doses.Administered.Janssen,
    unknown = doses.Administered.Unknown
  ) %>%
  pivot_longer(!State, names_to = "vaccineName", values_to = "administeredCount")
dosesAdministered
```
```{r}
fivestatesdoses <- covid19Data1 %>%
  right_join(dosesAdministered, by = "State")  
fivestatesdoses
```

```{r}
dosesAdministeredregions <-
```


```{r}
ggplot(dosesAdministered, aes(x= vaccineName, weight = administeredCount)) + geom_bar() + facet_wrap(~State)
```
